Report on the Fourth Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE4)

### 2016-09-17 WSSSPE4_report/WSSSPE4_report.tex
more changes

\author{Daniel S.\ Katz$^{(1)}$,
Kyle E.\ Niemeyer$^{(2)}$,
}

@@

The Fourth Workshop on Sustainable Software for Science: Practice and Experiences
(WSSSPE4)\footnote{\url{http://wssspe.researchcomputing.org.uk/wssspe4/}} was
held on 12--14 September 2016 in Manchester, England, UK.
This location and date was selected so that WSSSPE4 immediately preceded the
First Research Software Engineers (RSE) Conference.
Previous events in the WSSSPE series include
WSSSPE1\footnote{\url{http://wssspe.researchcomputing.org.uk/wssspe1/}}~\cite{WSSSPE1-pre-report,WSSSPE1},
held in conjunction with SC13;
WSSSPE1.1\footnote{\url{http://wssspe.researchcomputing.org.uk/wssspe1-1/}}, a
focused workshop organized jointly with the SciPy
conference\footnote{\url{https://conference.scipy.org/scipy2014/participate/wssspe/}};
WSSSPE2\footnote{\url{http://wssspe.researchcomputing.org.uk/wssspe2/}}~\cite{WSSSPE2-pre-report,WSSSPE2},
held in conjunction with SC14;
WSSSPE2.1\footnote{\url{http://wssspe.researchcomputing.org.uk/wssspe2-1/}}, a
focused workshop organized again jointly with
SciPy\footnote{\url{http://scipy2015.scipy.org/ehome/115969/286469/}};
and WSSSPE3\footnote{\url{http://wssspe.researchcomputing.org.uk/wssspe3/}}~\cite{WSSSPE3},
held in Boulder, Colorado, USA.
Note that the WSSSPE series changed its name from ``Workshop on Sustainable
Software for Science: Practice and Experiences'' to ``Working towards
Sustainable Software for Science: Practice and Experiences'' to better reflect
the movement towards a community in addition to the associated workshop series.\katznote{really?  I must have missed that.  I remember the first workshop as being `working towards' and then future workshops as being `workshop on'.}

@@

The CoC was introduced at the beginning of WSSSPE4, with the CoC subcommittee
and a general email address introduced for reporting concerns or incidents, or
asking questions.
\todo{mention any/lack of violations/reports?}

This report is based on
\todo{what report is based on}

@@

WSSSPE4 was based on the work done in WSSSPE1, WSSSPE2, and WSSSPE3, but aimed
at \todo{what aimed at?}

two tracks:

@@

\todo{update}
The calls for participation requested papers describing lightning talks,
where each author could make a brief statement about work that either had been
done or was needed;
idea papers, which present implementable proposals related to Track 1;
position papers, which are longer, not previously published papers related to
Track 2 specifically discussing what we can do to improve sustainable scientific
software in the short term, starting with where we are today;
experience papers, longer papers related to Track 2 that discuss current
practices and experiences and how they have been used to improve the quality of
today’s research software and/or the experiences of its developers;
and demos, brief papers describing a 10--15 minute demonstration of a tool or
process relevant to Track 2 that improves the quality of today's research
software and\slash or the experiences of its developers.

There were~24 lightning talks submitted; after a peer-review process, 16
were accepted, as discussed further in Section~\ref{sec:lightning}.
\todo{update}

@@

@@

@@

Intro and description
%
\begin{enumerate}
\item \textbf{Names: \emph{title}}.
Description

@@

\todo{description of panel discussion on best practices}

@@

\todo{go through subsections that have both content here and appendix, and add pointer from text in the subsection here to the corresponding appendix}

@@

\todo{discuss survey responses}

>>>>

\author{Daniel S.\ Katz$^{(1)}$,
Kyle E.\ Niemeyer$^{(2)}$, 
Author order to be determined based on contributions
}

@@

The Fourth Workshop on Sustainable Software for Science: Practice and Experiences
(WSSSPE4)\footnote{\url{http://wssspe.researchcomputing.org.uk/wssspe4/}} was
held over 2 1/2 days on 12--14 September 2016 in Manchester, England, UK.
This location and date was selected so that WSSSPE4 immediately preceded the
First Research Software Engineers (RSE) Conference.
Previous events in the WSSSPE series include
WSSSPE1\footnote{\url{http://wssspe.researchcomputing.org.uk/wssspe1/}}~\cite{WSSSPE1-pre-report,WSSSPE1},
held in conjunction with SC13;
WSSSPE1.1\footnote{\url{http://wssspe.researchcomputing.org.uk/wssspe1-1/}}, a
focused workshop organized jointly with the SciPy
conference\footnote{\url{https://conference.scipy.org/scipy2014/participate/wssspe/}};
WSSSPE2\footnote{\url{http://wssspe.researchcomputing.org.uk/wssspe2/}}~\cite{WSSSPE2-pre-report,WSSSPE2},
held in conjunction with SC14;
WSSSPE2.1\footnote{\url{http://wssspe.researchcomputing.org.uk/wssspe2-1/}}, a
focused workshop organized again jointly with
SciPy\footnote{\url{http://scipy2015.scipy.org/ehome/115969/286469/}};
and WSSSPE3\footnote{\url{http://wssspe.researchcomputing.org.uk/wssspe3/}}~\cite{WSSSPE3},
held in Boulder, Colorado, USA.
Note that the first WSSSPE workshop was named
``Working towards
Sustainable Software for Science: Practice and Experiences,'' which remains the meaning
of the WSSSPE group, but the workshops after that were named 
``Workshop on Sustainable
Software for Science: Practice and Experiences.'' Together these reflect
that WSSSPE is both a community and a set of workshops.

@@

The CoC was introduced at the beginning of WSSSPE4, with the CoC subcommittee
and a general email address introduced for reporting concerns or incidents, or
asking questions.  There was one concern mentioned to the CoC subcommittee
after the first half day,
regarding how part of the workshop was being run, and we changed the workshop
to address this.

This report is based on the events at the workshop and the submitted materials.
There events included discussion of the WSSSPE mission and vision, a keynote,
a set of presentations, a panel, and a number of working groups.  About a day of the
workshop was spent with participants in the working groups, which occurred in parallel
with each other.

@@

WSSSPE4 was based on the work done in WSSSPE1, WSSSPE2, and WSSSPE3, but aimed
at producing working groups that better continued working after the workshop ended.
In addition, based on feedback after WSSSPE3, it became clear that WSSSPE attendees
had two different motivations in participating.  One motivation was to make a better future
for research software, and the other was to immediately do better research software development.
This led to the idea of WSSSPE4 being partially divided into two tracks:

@@

The call for participation requested 
idea papers, which present implementable proposals related to Track 1;
position papers, which are longer, not previously published papers related to
Track 2 specifically discussing what we can do to improve sustainable scientific
software in the short term, starting with where we are today;
experience papers, longer papers related to Track 2 that discuss current
practices and experiences and how they have been used to improve the quality of
today's research software and/or the experiences of its developers;
demos, brief papers describing a tool or
process that would be demonstrated, relevant to Track 2 that improves the quality of today's research
software and\slash or the experiences of its developers; and
extended abstracts describing lightning talks,
where each author could make a brief statement about work that either had been
done or was needed.

There were~24 lightning talks submitted; after a peer-review process, 16
were accepted, as discussed further in Section~\ref{sec:lightning}.
\todo{update, add paper stats too}

@@

\todo{need to mention/link to the proceedings somewhere, though maybe not here.}

@@

\note{need a person to write this}

The keynote was given by Patricia Lago and entitled ``\todo{title}''.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Mission and vision}\label{sec:mission}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Papers and demos} \label{sec:papers}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\note{need a person to write this}

Intro and description \katznote{not sure enum list is right here}
%
\begin{enumerate}
\item \textbf{Authors: \emph{title}}.
Description

@@

\note{need a person to write this}

Intro and description \katznote{not sure enum list is right here}
%
\begin{enumerate}
\item \textbf{Authors: \emph{title}}.
Description

@@

\todo{description of panel discussion on best practices} \note{asked Simon to write}

@@

\todo{describe process for assembling people into working groups: vision, gaps, projects exercise, then soapboxes}

\todo{go through subsections that have both content here and appendix, and add pointer from text in the subsection here to the corresponding appendix} \todo{though think about if an appendix is the right thing if we decide to publish in JORS again}

@@

\note{need a person to write this - Dan asked Lorraine}

\todo{discuss survey responses; maybe put detailed responses in an appendix}
